---
title: "Snake_venom_analysis"
author: "Kevin Healy"
date: "15 Jan  2018"
output:
  html_document: default
  pdf_document: default
---


# Upload stuff
## Packages

First the required packages are uplaoded. These include .MCMCglmm' which is used for the main linear analysis, 'phytools' and 'caper'. to handle the phylogenies and finaly 'hdrcde', 'wesanderson' and the 'MultiDisPlot.R' function to make some of the plots.

```{r packages}

library(phytools)
library(caper)
library(MCMCglmm)
library(wesanderson)
library(hdrcde)
source("MultiDisPlot.R")
```


## data
Next we uplaod the main data file. See the methods section in the main paper for a description of the data.

```{r data}
ld50_data <- read.csv("S2_snake_data.csv",header=T,sep=",")

```

## Phylogeny
We use the phylogeny from Pyron et al 2014 to control for phylogeny in our main model.

```{r phylo}
tree <- read.tree("liz_and_snake_time_tree.txt")
#this seems to fix a dublicate problem caused by polynomies in the tree
Tree<-makeLabel(tree)
Tree<-makeLabel(Tree)
```


## Data transformation
To allow for comparison of effect sizes we log transform all continous variables and z-score them.

```{r transform}

ld50_clean  <- ld50_data
#length
len.v <- log10(ld50_clean$length_cm)
len.v <- (len.v - mean(len.v))/(sd(len.v))

#mass
mass <- log10(ld50_clean $mass_grams)
mass <- (mass- mean(mass))/(sd(mass))

animal <- ld50_clean $species
species <- ld50_clean $species

##volume
vol.v <- log10(as.numeric(as.vector(ld50_clean $venom_yield_mg)))
vol.v <- (vol.v - mean(vol.v))/(sd(vol.v))

#dimension
dimensions <- ld50_clean $dimensions

#LD50
ld50_10 <- log10(as.vector(ld50_clean $ld50_mg))
ld50_mg <- (ld50_10- mean(ld50_10))/(sd(ld50_10))


ld50_method <- ld50_clean $ld50_method

eggs <- ld50_clean[,c("eggs")] 
egg.bin <- rep(0,length(eggs))
for(i in 1:(length(eggs))){
	
if(ld50_clean[i,c("eggs")] > 0) egg.bin[i] <- "yes"
else egg.bin[i] <- "no"
}

weigted.dist.v <- (ld50_clean$phylo_distance_diet_model_my)
weigted.dist.v <- (weigted.dist.v- mean(weigted.dist.v))/(sd(weigted.dist.v))

env <- ld50_clean$environment

con <- ld50_clean$constriction

fam <- ld50_clean$family


mod_class <- ld50_clean$ld50_model_class

```


## full model

Now we set up the model to insure that there is a corresponding animal column within the dataframe in order for the phylogeny to match up. We also include as speerate "species" term to account for variation assocaited with multiple measures for a single species.

```{r full model setup}


###full model variables
mydat.max <- data.frame(animal, 
                        species, 
                        mass, 
                        weigted.dist.v = weigted.dist.v, 
                        vol.v, 
                        dimensions = dimensions, 
                        ld50 = ld50_10, 
                        ld50_method, 
                        dimensions, 
                        egg.bin, 
                        env, 
                        con, 
                        fam)

spec_dropped <-as.data.frame(matrix(0, nrow = c(length(levels(mydat.max[,"species"]))), 
                                    ncol = c(2), dimnames = list(c(),c("species","dummy"))))
spec_dropped[,1] <- levels(mydat.max$species)
spec_dropped[,2] <- rep(0,length = c(length(levels(mydat.max[,"species"]))))
spec_dropped <- as.data.frame(spec_dropped)

max.dat <- comparative.data(data= spec_dropped, 
                            phy= Tree, 
                            names.col="species", 
                            vcv=FALSE)

max.tree <- chronoMPL(max.dat$phy)

max.dropped <- max.dat$dropped$unmatched.rows

mydat.MAX <- mydat.max
if((length(max.dropped) > c(0)) == TRUE){
for(i in 1:(length(max.dropped))){
mydat.MAX <-mydat.MAX[mydat.MAX$species != max.dropped[i],]
}
}


##clean up the final matched data.frame
mydat.MAX <- as.data.frame(as.matrix(mydat.MAX))

mydat.MAX <- data.frame(animal = mydat.MAX$species, 
species = mydat.MAX$species, 
mass.aa = (as.numeric(as.character(mydat.MAX$mass))), 
weigted.dist.v = (as.numeric(as.character(mydat.MAX$weigted.dist.v))), 
vol.v = (as.numeric(as.character(mydat.MAX$vol.v))), 
ld50 = (as.numeric(as.character(mydat.MAX$ld50))), 
ld50_method  = factor(mydat.MAX$ld50_method, 
                       levels = c("sc","iv","ip","im")), 
dim = mydat.MAX$dimensions, 
egg.bin = mydat.MAX$egg.bin, 
env = factor(mydat.MAX$env, 
             levels = c("terrestrial","arboreal","aquatic")),
con = mydat.MAX$con,
fam = mydat.MAX$fam)



```



##Set the parameters
Here we set the number of iterations, the thinning and the burn-in

```{r parameters}
#if you want to include species and ld50 method as a fixed term.

#nitt <- 1200000
#thin <- 500
#burnin <- 200000

nitt <- 12000
thin <- 5
burnin <- 2000

```


##Set the prior

Set the proir. Here we stick with a x prior.

```{r prior}
prior<-list(R = list(V = diag(2), nu=0.002), 
            G = list(G1=list(V = diag(2)/2,n = 2, 
                             alpha.mu=rep(0,2), 
                             alpha.V=diag(2)*10^3),
                     G2=list(V = diag(2)/2,n = 2, 
                             alpha.mu=rep(0,2), 
                             alpha.V=diag(2)*10^3)))
```

##Run the full model
Now we run the full model with both venom volume and Ld50 included as response variables. We also allow for co-variance between these two terms.

```{r full model}

full_model <- MCMCglmm(cbind(vol.v,ld50) ~ 
                         trait:mass.aa + 
                         trait:ld50_method + 
                         trait:dim + 
                         trait:egg.bin + 
                         trait:weigted.dist.v, 
                       random = ~us(trait):animal + 
                                 us(trait):species,
                       rcov=~us(trait):units, 
                       data = mydat.MAX, 
                       pedigree = max.tree, 
                       prior = prior,
                       family=c("gaussian","gaussian"), 
                       nitt = nitt, thin = thin, burnin = burnin,
                       verbose= FALSE)

summary(full_model)
```


##Run the full model
Now we calulated the proportion of variance associated with phylogeny, species level and risidual for both volume and LD50

```{r H2}
ld_50_prop_phlyo <- full_model$VCV[,"traitld50:traitld50.animal"]/
                    (full_model$VCV[,"traitld50:traitld50.animal"] +
                     full_model$VCV[,"traitld50:traitld50.species"] +
                     full_model$VCV[,"traitld50:traitld50.units"] +
                         full_model$VCV[,"traitld50:traitvol.v.animal"] + 
                         full_model$VCV[,"traitld50:traitvol.v.species"] + 
                         full_model$VCV[,"traitld50:traitvol.v.units"])

ld_50_prop_species <- full_model$VCV[,"traitld50:traitld50.species"]/
                    (full_model$VCV[,"traitld50:traitld50.animal"] +
                     full_model$VCV[,"traitld50:traitld50.species"] +
                     full_model$VCV[,"traitld50:traitld50.units"]+
                         full_model$VCV[,"traitld50:traitvol.v.animal"] + 
                         full_model$VCV[,"traitld50:traitvol.v.species"] + 
                         full_model$VCV[,"traitld50:traitvol.v.units"])

ld_50_prop_units <- full_model$VCV[,"traitld50:traitld50.units"]/
                    (full_model$VCV[,"traitld50:traitld50.animal"] +
                     full_model$VCV[,"traitld50:traitld50.species"] +
                     full_model$VCV[,"traitld50:traitld50.units"]+
                         full_model$VCV[,"traitld50:traitvol.v.animal"] + 
                         full_model$VCV[,"traitld50:traitvol.v.species"] + 
                         full_model$VCV[,"traitld50:traitvol.v.units"])

hdr(ld_50_prop_phlyo)
hdr(ld_50_prop_species)
hdr(ld_50_prop_units)



vol_prop_phlyo <-  full_model$VCV[,"traitvol.v:traitvol.v.animal"]/
                    (full_model$VCV[,"traitvol.v:traitvol.v.animal"] +
                     full_model$VCV[,"traitvol.v:traitvol.v.species"] +
                     full_model$VCV[,"traitvol.v:traitvol.v.units"] +
                         full_model$VCV[,"traitld50:traitvol.v.animal"] + 
                         full_model$VCV[,"traitld50:traitvol.v.species"] + 
                         full_model$VCV[,"traitld50:traitvol.v.units"])

vol_prop_spec <-  full_model$VCV[,"traitld50:traitvol.v.species"]/
                    (full_model$VCV[,"traitvol.v:traitvol.v.animal"] +
                     full_model$VCV[,"traitvol.v:traitvol.v.species"] +
                     full_model$VCV[,"traitvol.v:traitvol.v.units"] +
                         full_model$VCV[,"traitld50:traitvol.v.animal"] + 
                         full_model$VCV[,"traitld50:traitvol.v.species"] + 
                         full_model$VCV[,"traitld50:traitvol.v.units"])

vol_prop_units <-  full_model$VCV[,"traitvol.v:traitvol.v.units"]/
                    (full_model$VCV[,"traitvol.v:traitvol.v.animal"] +
                     full_model$VCV[,"traitvol.v:traitvol.v.species"] +
                     full_model$VCV[,"traitvol.v:traitvol.v.units"] +
                         full_model$VCV[,"traitld50:traitvol.v.animal"] + 
                         full_model$VCV[,"traitld50:traitvol.v.species"] + 
                         full_model$VCV[,"traitld50:traitvol.v.units"])


hdr(vol_prop_phlyo)
hdr(vol_prop_spec)
hdr(vol_prop_units)

```

###plot the full model

We plot the posterior distributions in a table using the `MultiDisPlot` function.

```{r MultiDisPlot plot}

vol_post <- list(full_model$Sol[,1],
                 full_model$Sol[,"traitvol.v:mass.aa"],
                 full_model$Sol[,"traitvol.v:ld50_methodiv"],
                 full_model$Sol[,"traitvol.v:ld50_methodip"],
                 full_model$Sol[,"traitvol.v:ld50_methodim"],
                 full_model$Sol[,"traitvol.v:dim3"],
                 full_model$Sol[,"traitvol.v:egg.binyes"],
                 full_model$Sol[,"traitvol.v:weigted.dist.v"],
                 vol_prop_phlyo,
                 full_model$VCV[,"traitld50:traitvol.v.animal"],
                 vol_prop_spec,
                 full_model$VCV[,"traitld50:traitvol.v.species"],
                 vol_prop_units,
                 full_model$VCV[,"traitld50:traitvol.v.units"])
MultiDisPlot(vol_post)
abline(v = 0, lty = 2)

ld50_post <- list(full_model$Sol[,1],
                 full_model$Sol[,"traitld50:mass.aa"],
                 full_model$Sol[,"traitld50:ld50_methodiv"],
                 full_model$Sol[,"traitld50:ld50_methodip"],
                 full_model$Sol[,"traitld50:ld50_methodim"],
                 full_model$Sol[,"traitld50:dim3"],
                 full_model$Sol[,"traitld50:egg.binyes"],
                 full_model$Sol[,"traitld50:weigted.dist.v"],
                 ld_50_prop_phlyo,
                 full_model$VCV[,"traitld50:traitvol.v.animal"],
                 ld_50_prop_species,
                 full_model$VCV[,"traitld50:traitvol.v.species"],
                 ld_50_prop_units,
                 full_model$VCV[,"traitld50:traitvol.v.units"]
                 )
MultiDisPlot(ld50_post)
abline(v = 0, lty = 2)
```


#Main figure Plots
Lets plot the main analysis
First the volume plot
```{r volume plot}
pal <- wes_palette("Zissou", 4, type = "continuous")


#pdf("vol_plot.pdf")
plot(mydat.MAX[mydat.MAX$dim == "1" & mydat.MAX$egg.bin == "no", "vol.v"] ~  mydat.MAX[mydat.MAX$dim == "1" & mydat.MAX$egg.bin == "no", "mass.aa"], 
     pch = 16, 
     xlab= expression(log[10]*Mass*(g)), 
     ylab = expression(log[10]*venom*(mg)), 
     bty = "l" , 
     col = pal[4], 
     ylim = c(-1,3.5),  
     xlim = c(1.5,4.5)) 

points(mydat.MAX[mydat.MAX$dim == "3"  & mydat.MAX$egg.bin == "no", "vol.v"] ~ mydat.MAX[mydat.MAX$dim == "3"  & mydat.MAX$egg.bin == "no" , "mass.aa"], 
       pch = 16, 
       col = pal[1])

points(mydat.MAX[mydat.MAX$dim == "1"  & mydat.MAX$egg.bin == "yes", "vol.v"] ~ mydat.MAX[mydat.MAX$dim == "1"  & mydat.MAX$egg.bin == "yes" , "mass.aa"], 
       pch = 17, 
       col = pal[4] )

points(mydat.MAX[mydat.MAX$dim == "3"  & mydat.MAX$egg.bin == "yes", "vol.v"] ~ mydat.MAX[mydat.MAX$dim == "3"  & mydat.MAX$egg.bin == "yes" , "mass.aa"], 
       pch = 17, 
       col = pal[1] )


points(mydat.MAX[ mydat.MAX$species ==  "Thamnophis_elegans", "vol.v"] ~
       mydat.MAX[ mydat.MAX$species ==  "Thamnophis_elegans", "mass.aa"] 
       , col = "white", 
       pch = 20)

points(mydat.MAX[ mydat.MAX$species ==  "Emydocephalus_annulatus", "vol.v"] ~
       mydat.MAX[ mydat.MAX$species ==  "Emydocephalus_annulatus", "mass.aa"] 
       , col = "white", 
       pch = 20)

points(mydat.MAX[ mydat.MAX$species ==  "Causus_rhombeatus", "vol.v"] ~
       mydat.MAX[ mydat.MAX$species ==  "Causus_rhombeatus", "mass.aa"] 
       , col = "white", 
       pch = 20)

points(mydat.MAX[ mydat.MAX$species ==  "Atractaspis_bibronii", "vol.v"] ~
       mydat.MAX[ mydat.MAX$species ==  "Atractaspis_bibronii", "mass.aa"] 
       , col = "white", 
       pch = 20)

points(mydat.MAX[ mydat.MAX$species ==  "Hydrophis_elegans", "vol.v"] ~
       mydat.MAX[ mydat.MAX$species ==  "Hydrophis_elegans", "mass.aa"] 
       , col = "white", 
       pch = 20)

points(mydat.MAX[ mydat.MAX$species ==  "Agkistrodon_piscivorus", "vol.v"] ~
       mydat.MAX[ mydat.MAX$species ==  "Agkistrodon_piscivorus", "mass.aa"] 
       , col = "white", 
       pch = 20)

points(mydat.MAX[ mydat.MAX$species ==  "Ophiophagus_hannah", "vol.v"] ~
       mydat.MAX[ mydat.MAX$species ==  "Ophiophagus_hannah", "mass.aa"] 
       , col = "white", 
       pch = 20)

points(mydat.MAX[ mydat.MAX$species ==  "Daboia_russelii", "vol.v"] ~
       mydat.MAX[ mydat.MAX$species ==  "Daboia_russelii", "mass.aa"] 
       , col = "white", 
       pch = 20)

points(mydat.MAX[ mydat.MAX$species ==  "Bungarus_multicinctus", "vol.v"] ~
       mydat.MAX[ mydat.MAX$species ==  "Bungarus_multicinctus", "mass.aa"] 
       , col = "white", 
       pch = 20)

points(mydat.MAX[ mydat.MAX$species ==  "Oxyuranus_scutellatus", "vol.v"] ~
       mydat.MAX[ mydat.MAX$species ==  "Oxyuranus_scutellatus", "mass.aa"] 
       , col = "white", 
       pch = 20)

abline(summary(full_model)$solutions[1,1], 
       summary(full_model)$solutions[2,1], 
       col = pal[4], 
       lw =2)

abline(summary(full_model)$solutions[1,1] + summary(full_model)$solutions[10,1],
       summary(full_model)$solutions[2,1], 
       col = pal[1], 
       lw =2 )

#dev.off()
```

nest the LD50 plot.

```{r ld50 plot}
###plots
###LD50

#pdf("ld50_plot.pdf")

plot(mydat.MAX$ld50 ~ (mydat.MAX$weigted.dist.v), 
     pch = 16, 
     xlab=  expression(paste("Mean phylogenetic distance between diet and ", 
     LD[50], 
     " model (Myr)", 
     sep= " ")), 
     ylab = expression(paste(log[10]*LD[50], 
     " (mg/kg)")), 
     bty = "l", 
     col = pal[3])

points(mydat.MAX[mydat.MAX$ld50_method == "iv", "ld50"] ~ 
         mydat.MAX[mydat.MAX$ld50_method == "iv", "weigted.dist.v"], 
       pch = 16, 
       col = pal[4])

points(mydat.MAX[mydat.MAX$ld50_method == "sc", "ld50"] ~ 
         mydat.MAX[mydat.MAX$ld50_method == "sc" , "weigted.dist.v"], 
       pch = 16, 
       col = pal[3])

points(mydat.MAX[mydat.MAX$ld50_method == "im", "ld50"] ~ 
         mydat.MAX[mydat.MAX$ld50_method == "im" , "weigted.dist.v"], 
       pch = 16, 
       col = pal[1])

points(mydat.MAX[mydat.MAX$ld50_method == "ip", "ld50"] ~ 
         mydat.MAX[mydat.MAX$ld50_method == "ip" , "weigted.dist.v"], 
       pch = 16, 
       col = pal[2])

points(mydat.MAX[mydat.MAX$egg.bin == "yes", "ld50"] ~ 
         mydat.MAX[mydat.MAX$egg.bin == "yes" , "weigted.dist.v"], 
       pch = 17, 
       col = pal[3])


points(mydat.MAX[ mydat.MAX$species ==  "Thamnophis_elegans", "ld50"] ~
       mydat.MAX[ mydat.MAX$species ==  "Thamnophis_elegans", "weigted.dist.v"] 
       , col = "white", 
       pch = 20)

points(mydat.MAX[ mydat.MAX$species ==  "Emydocephalus_annulatus", "ld50"] ~
       mydat.MAX[ mydat.MAX$species ==  "Emydocephalus_annulatus", "weigted.dist.v"] 
       , col = "white", 
       pch = 20)

points(mydat.MAX[ mydat.MAX$species ==  "Causus_rhombeatus", "ld50"] ~
       mydat.MAX[ mydat.MAX$species ==  "Causus_rhombeatus", "weigted.dist.v"] 
       , col = "white", 
       pch = 20)


points(mydat.MAX[ mydat.MAX$species ==  "Agkistrodon_piscivorus", "ld50"] ~
       mydat.MAX[ mydat.MAX$species ==  "Agkistrodon_piscivorus", "weigted.dist.v"] 
       , col = "white", 
       pch = 20)

points(mydat.MAX[ mydat.MAX$species ==  "Ophiophagus_hannah", "ld50"] ~
       mydat.MAX[ mydat.MAX$species ==  "Ophiophagus_hannah", "weigted.dist.v"] 
       , col = "white", 
       pch = 20)

points(mydat.MAX[ mydat.MAX$species ==  "Daboia_russelii", "ld50"] ~
       mydat.MAX[ mydat.MAX$species ==  "Daboia_russelii", "weigted.dist.v"] 
       , col = "white", 
       pch = 20)

points(mydat.MAX[ mydat.MAX$species ==  "Bungarus_multicinctus", "ld50"] ~
       mydat.MAX[ mydat.MAX$species ==  "Bungarus_multicinctus", "weigted.dist.v"] 
       , col = "white", 
       pch = 20)

points(mydat.MAX[ mydat.MAX$species ==  "Oxyuranus_scutellatus", "ld50"] ~
       mydat.MAX[ mydat.MAX$species ==  "Oxyuranus_scutellatus", "weigted.dist.v"] 
       , col = "white", 
       pch = 20)


abline(summary(full_model)$solutions[1,1], 
       summary(full_model)$solutions[15,1], 
       col = pal[2], 
       lw =2)

#dev.off()

```


#plot phylo
We can also plot the phylogeny

```{r phylo plot}

#First make a named vector of the mean trait value.

plot_LD50 <- vector()
plot_vol <- vector()


for(i in 1:length(levels(species))){
  
  plot_LD50[i] <-  min(ld50_data[ ld50_data$species ==  levels(species)[i] ,
                                  "ld50_mg"])
  plot_vol[i] <-  (ld50_data[ ld50_data$species ==  levels(species)[i] ,
                              "venom_yield_mg"])

}

names(plot_LD50) <- sort(max.tree$tip.label)
names(plot_vol) <- sort(max.tree$tip.label)

#pdf("phylobar_ld50.pdf")
plotTree.wBars(tree = max.tree, 
               x = plot_LD50, 
               scale= 1, 
               type="fan", 
               cex = 0.2, 
               tip.labels = T, 
               fsize=0.37, 
               width = 1)
#dev.off()

#pdf("phylobar_vol.pdf")
plotTree.wBars(tree = max.tree, 
               x = plot_vol, 
               scale= 0.04, 
               type="fan", 
               cex = 0.2, 
               tip.labels = T, 
               fsize=0.37, 
               width = 1)
#dev.off()

```




#Supplemntary models

Now lets rerun the analysis for when we include constriction, family level and prey size.

##Run the constriction model
Constriction Model

```{r constriction model}
#if you want to include species and ld50 method as a fixed term.
prior<-list(R = list(V = diag(2), nu=0.002), 
            G = list(G1=list(V = diag(2)/2,n = 2, 
                             alpha.mu=rep(0,2), 
                             alpha.V=diag(2)*10^3),
                     G2=list(V = diag(2)/2,n = 2, 
                             alpha.mu=rep(0,2), 
                             alpha.V=diag(2)*10^3)))


######for constriction
con_mod <- MCMCglmm(cbind(vol.v,ld50) ~ 
                      trait:mass.aa + 
                      trait:ld50_method+ 
                      trait:con + 
                      trait:dim + 
                      trait:egg.bin + 
                      trait:weigted.dist.v, 
                    random = ~us(trait):animal + 
                              us(trait):species,
                    rcov=~us(trait):units, 
                    data = mydat.MAX, 
                    pedigree = max.tree, 
                    prior = prior,
                    family=c("gaussian","gaussian"), 
                    nitt = nitt, thin = thin, burnin = burnin,
                    verbose = FALSE)


summary(con_mod)


ld_50_con_phlyo <- con_mod$VCV[,"traitld50:traitld50.animal"]/
                    (con_mod$VCV[,"traitld50:traitld50.animal"] +
                     con_mod$VCV[,"traitld50:traitld50.species"] +
                     con_mod$VCV[,"traitld50:traitld50.units"] +
                         con_mod$VCV[,"traitld50:traitvol.v.animal"] + 
                         con_mod$VCV[,"traitld50:traitvol.v.species"] + 
                         con_mod$VCV[,"traitld50:traitvol.v.units"])

ld_50_con_species <- con_mod$VCV[,"traitld50:traitld50.species"]/
                    (con_mod$VCV[,"traitld50:traitld50.animal"] +
                     con_mod$VCV[,"traitld50:traitld50.species"] +
                     con_mod$VCV[,"traitld50:traitld50.units"]+
                         con_mod$VCV[,"traitld50:traitvol.v.animal"] + 
                         con_mod$VCV[,"traitld50:traitvol.v.species"] + 
                         con_mod$VCV[,"traitld50:traitvol.v.units"])

ld_50_con_units <- con_mod$VCV[,"traitld50:traitld50.units"]/
                    (con_mod$VCV[,"traitld50:traitld50.animal"] +
                     con_mod$VCV[,"traitld50:traitld50.species"] +
                     con_mod$VCV[,"traitld50:traitld50.units"]+
                         con_mod$VCV[,"traitld50:traitvol.v.animal"] + 
                         con_mod$VCV[,"traitld50:traitvol.v.species"] + 
                         con_mod$VCV[,"traitld50:traitvol.v.units"])

hdr(ld_50_con_phlyo)
hdr(ld_50_con_species)
hdr(ld_50_con_units)



vol_con_phlyo <-  con_mod$VCV[,"traitvol.v:traitvol.v.animal"]/
                    (con_mod$VCV[,"traitvol.v:traitvol.v.animal"] +
                     con_mod$VCV[,"traitvol.v:traitvol.v.species"] +
                     con_mod$VCV[,"traitvol.v:traitvol.v.units"] +
                         con_mod$VCV[,"traitld50:traitvol.v.animal"] + 
                         con_mod$VCV[,"traitld50:traitvol.v.species"] + 
                         con_mod$VCV[,"traitld50:traitvol.v.units"])

vol_con_spec <-  con_mod$VCV[,"traitld50:traitvol.v.species"]/
                    (con_mod$VCV[,"traitvol.v:traitvol.v.animal"] +
                     con_mod$VCV[,"traitvol.v:traitvol.v.species"] +
                     con_mod$VCV[,"traitvol.v:traitvol.v.units"] +
                         con_mod$VCV[,"traitld50:traitvol.v.animal"] + 
                         con_mod$VCV[,"traitld50:traitvol.v.species"] + 
                         con_mod$VCV[,"traitld50:traitvol.v.units"])

vol_con_units <-  con_mod$VCV[,"traitvol.v:traitvol.v.units"]/
                    (con_mod$VCV[,"traitvol.v:traitvol.v.animal"] +
                     con_mod$VCV[,"traitvol.v:traitvol.v.species"] +
                     con_mod$VCV[,"traitvol.v:traitvol.v.units"] +
                         con_mod$VCV[,"traitld50:traitvol.v.animal"] + 
                         con_mod$VCV[,"traitld50:traitvol.v.species"] + 
                         con_mod$VCV[,"traitld50:traitvol.v.units"])


hdr(vol_con_phlyo)
hdr(vol_con_spec)
hdr(vol_con_units)

```


##Run the environment model
Model with environment type included

```{r env model}
#if you want to include species and ld50 method as a fixed term.
prior<-list(R = list(V = diag(2), nu=0.002), 
            G = list(G1=list(V = diag(2)/2,n = 2, 
                             alpha.mu=rep(0,2), 
                             alpha.V=diag(2)*10^3),
                     G2=list(V = diag(2)/2,n = 2, 
                             alpha.mu=rep(0,2), 
                             alpha.V=diag(2)*10^3)))

######for environment
env_mod <- MCMCglmm(cbind(vol.v,ld50) ~ 
                      trait:mass.aa + 
                      trait:ld50_method + 
                      trait:env + 
                      trait:egg.bin + 
                      trait:weigted.dist.v, 
                    random = ~us(trait):animal + 
                              us(trait):species,
                    rcov=~us(trait):units, 
                    data = mydat.MAX, 
                    pedigree = max.tree, 
                    prior = prior,
                    family=c("gaussian","gaussian"), 
                    nitt = nitt, thin =thin, burnin = burnin,
                    verbose = FALSE)

summary(env_mod)


ld_50_env_phlyo <- env_mod$VCV[,"traitld50:traitld50.animal"]/
                    (env_mod$VCV[,"traitld50:traitld50.animal"] +
                     env_mod$VCV[,"traitld50:traitld50.species"] +
                     env_mod$VCV[,"traitld50:traitld50.units"] +
                         env_mod$VCV[,"traitld50:traitvol.v.animal"] + 
                         env_mod$VCV[,"traitld50:traitvol.v.species"] + 
                         env_mod$VCV[,"traitld50:traitvol.v.units"])

ld_50_env_species <- env_mod$VCV[,"traitld50:traitld50.species"]/
                    (env_mod$VCV[,"traitld50:traitld50.animal"] +
                     env_mod$VCV[,"traitld50:traitld50.species"] +
                     env_mod$VCV[,"traitld50:traitld50.units"]+
                         env_mod$VCV[,"traitld50:traitvol.v.animal"] + 
                         env_mod$VCV[,"traitld50:traitvol.v.species"] + 
                         env_mod$VCV[,"traitld50:traitvol.v.units"])

ld_50_env_units <- env_mod$VCV[,"traitld50:traitld50.units"]/
                    (env_mod$VCV[,"traitld50:traitld50.animal"] +
                     env_mod$VCV[,"traitld50:traitld50.species"] +
                     env_mod$VCV[,"traitld50:traitld50.units"]+
                         env_mod$VCV[,"traitld50:traitvol.v.animal"] + 
                         env_mod$VCV[,"traitld50:traitvol.v.species"] + 
                         env_mod$VCV[,"traitld50:traitvol.v.units"])

hdr(ld_50_env_phlyo)
hdr(ld_50_env_species)
hdr(ld_50_env_units)



vol_env_phlyo <-  env_mod$VCV[,"traitvol.v:traitvol.v.animal"]/
                    (env_mod$VCV[,"traitvol.v:traitvol.v.animal"] +
                     env_mod$VCV[,"traitvol.v:traitvol.v.species"] +
                     env_mod$VCV[,"traitvol.v:traitvol.v.units"] +
                         env_mod$VCV[,"traitld50:traitvol.v.animal"] + 
                         env_mod$VCV[,"traitld50:traitvol.v.species"] + 
                         env_mod$VCV[,"traitld50:traitvol.v.units"])

vol_env_spec <-  env_mod$VCV[,"traitld50:traitvol.v.species"]/
                    (env_mod$VCV[,"traitvol.v:traitvol.v.animal"] +
                     env_mod$VCV[,"traitvol.v:traitvol.v.species"] +
                     env_mod$VCV[,"traitvol.v:traitvol.v.units"] +
                         env_mod$VCV[,"traitld50:traitvol.v.animal"] + 
                         env_mod$VCV[,"traitld50:traitvol.v.species"] + 
                         env_mod$VCV[,"traitld50:traitvol.v.units"])

vol_env_units <-  env_mod$VCV[,"traitvol.v:traitvol.v.units"]/
                    (env_mod$VCV[,"traitvol.v:traitvol.v.animal"] +
                     env_mod$VCV[,"traitvol.v:traitvol.v.species"] +
                     env_mod$VCV[,"traitvol.v:traitvol.v.units"] +
                         env_mod$VCV[,"traitld50:traitvol.v.animal"] + 
                         env_mod$VCV[,"traitld50:traitvol.v.species"] + 
                         env_mod$VCV[,"traitld50:traitvol.v.units"])


hdr(vol_env_phlyo)
hdr(vol_env_spec)
hdr(vol_env_units)

```

##family model
Model with family level included

```{r fam model}
#if you want to include species and ld50 method as a fixed term.
prior<-list(R = list(V = diag(2), nu=0.002), 
            G = list(G1=list(V = diag(2)/2,n = 2, 
                             alpha.mu=rep(0,2), 
                             alpha.V=diag(2)*10^3),
                     G2=list(V = diag(2)/2,n = 2, 
                             alpha.mu=rep(0,2), 
                             alpha.V=diag(2)*10^3)))


######for constriction
fam_mod <- MCMCglmm(cbind(vol.v,ld50) ~ trait:mass.aa + 
                      trait:ld50_method + 
                      trait:fam + 
                      trait:dim + 
                      trait:egg.bin + 
                      trait:weigted.dist.v, 
                    random = ~us(trait):animal + 
                              us(trait):species,
                    rcov=~us(trait):units, 
                    data = mydat.MAX, 
                    pedigree = max.tree, 
                    prior = prior,
                    family=c("gaussian","gaussian"), 
                    nitt = nitt, thin =thin, burnin = burnin,
                    verbose = FALSE)


ld_50_fam_phlyo <- fam_mod$VCV[,"traitld50:traitld50.animal"]/
                    (fam_mod$VCV[,"traitld50:traitld50.animal"] +
                     fam_mod$VCV[,"traitld50:traitld50.species"] +
                     fam_mod$VCV[,"traitld50:traitld50.units"] +
                         fam_mod$VCV[,"traitld50:traitvol.v.animal"] + 
                         fam_mod$VCV[,"traitld50:traitvol.v.species"] + 
                         fam_mod$VCV[,"traitld50:traitvol.v.units"])

ld_50_fam_species <- fam_mod$VCV[,"traitld50:traitld50.species"]/
                    (fam_mod$VCV[,"traitld50:traitld50.animal"] +
                     fam_mod$VCV[,"traitld50:traitld50.species"] +
                     fam_mod$VCV[,"traitld50:traitld50.units"]+
                         fam_mod$VCV[,"traitld50:traitvol.v.animal"] + 
                         fam_mod$VCV[,"traitld50:traitvol.v.species"] + 
                         fam_mod$VCV[,"traitld50:traitvol.v.units"])

ld_50_fam_units <- fam_mod$VCV[,"traitld50:traitld50.units"]/
                    (fam_mod$VCV[,"traitld50:traitld50.animal"] +
                     fam_mod$VCV[,"traitld50:traitld50.species"] +
                     fam_mod$VCV[,"traitld50:traitld50.units"]+
                         fam_mod$VCV[,"traitld50:traitvol.v.animal"] + 
                         fam_mod$VCV[,"traitld50:traitvol.v.species"] + 
                         fam_mod$VCV[,"traitld50:traitvol.v.units"])

hdr(ld_50_fam_phlyo)
hdr(ld_50_fam_species)
hdr(ld_50_fam_units)



vol_fam_phlyo <-  fam_mod$VCV[,"traitvol.v:traitvol.v.animal"]/
                    (fam_mod$VCV[,"traitvol.v:traitvol.v.animal"] +
                     fam_mod$VCV[,"traitvol.v:traitvol.v.species"] +
                     fam_mod$VCV[,"traitvol.v:traitvol.v.units"] +
                         fam_mod$VCV[,"traitld50:traitvol.v.animal"] + 
                         fam_mod$VCV[,"traitld50:traitvol.v.species"] + 
                         fam_mod$VCV[,"traitld50:traitvol.v.units"])

vol_fam_spec <-  fam_mod$VCV[,"traitld50:traitvol.v.species"]/
                    (fam_mod$VCV[,"traitvol.v:traitvol.v.animal"] +
                     fam_mod$VCV[,"traitvol.v:traitvol.v.species"] +
                     fam_mod$VCV[,"traitvol.v:traitvol.v.units"] +
                         fam_mod$VCV[,"traitld50:traitvol.v.animal"] + 
                         fam_mod$VCV[,"traitld50:traitvol.v.species"] + 
                         fam_mod$VCV[,"traitld50:traitvol.v.units"])

vol_fam_units <-  fam_mod$VCV[,"traitvol.v:traitvol.v.units"]/
                    (fam_mod$VCV[,"traitvol.v:traitvol.v.animal"] +
                     fam_mod$VCV[,"traitvol.v:traitvol.v.species"] +
                     fam_mod$VCV[,"traitvol.v:traitvol.v.units"] +
                         fam_mod$VCV[,"traitld50:traitvol.v.animal"] + 
                         fam_mod$VCV[,"traitld50:traitvol.v.species"] + 
                         fam_mod$VCV[,"traitld50:traitvol.v.units"])


hdr(vol_fam_phlyo)
hdr(vol_fam_spec)
hdr(vol_fam_units)



summary(fam_mod)
```


#Models including prey size
Set up the dataset which includes prey and has a matching phylogeny

```{r prey model setup}


###full model
mydat.prey <- na.omit(data.frame(animal, 
                                 species, 
                                 mass, 
                                 weigted.dist.v = weigted.dist.v, 
                                 vol.v, 
                                 dimensions = dimensions, 
                                 ld50 = ld50_10, 
                                 ld50_method, 
                                 dimensions, 
                                 egg.bin, 
                                 env, 
                                 con, 
                                 fam, 
                                 prey_mass = ld50_clean$prey_size_g, 
                                 mod_class, 
                                 prey_max = ld50_clean$max_prey_g))


mydat.prey <- mydat.prey[mydat.prey$animal != "Causus_rhombeatus",]


spec_dropped <-as.data.frame(matrix(0, nrow = c(length(levels(mydat.prey[,"species"]))), 
                                    ncol = c(2), 
                                    dimnames = list(c(),c("species","dummy"))))

spec_dropped[,1] <- levels(mydat.prey$species)
spec_dropped[,2] <- rep(0,length = c(length(levels(mydat.prey[,"species"]))))
spec_dropped <- as.data.frame(spec_dropped)

prey.dat <- comparative.data(data= spec_dropped, 
                             phy= Tree, 
                             names.col="species", 
                             vcv=FALSE)

prey.tree <- chronoMPL(prey.dat$phy)

prey.dropped <- prey.dat$dropped$unmatched.rows

mydat.P <- mydat.prey
if((length(prey.dropped) > c(0)) == TRUE){
for(i in 1:(length(prey.dropped))){
mydat.P <-mydat.P[mydat.P$species != prey.dropped[i],]
}
}

mydat.P <- as.data.frame(as.matrix(mydat.P))

prey_mean <- (log10(as.numeric(as.character(mydat.P$prey_mass))) - mean(log10(as.numeric(as.character(mydat.P$prey_mass)))))/
  sd(log10(as.numeric(as.character(mydat.P$prey_mass))))


mydat.P <- data.frame(animal = mydat.P$species, 
species = mydat.P$species, 
mass.aa = (as.numeric(as.character(mydat.P$mass))), 
weigted.dist.v = (as.numeric(as.character(mydat.P$weigted.dist.v))), 
vol.v = (as.numeric(as.character(mydat.P$vol.v))), 
ld50 = (as.numeric(as.character(mydat.P$ld50))), 
ld50_method  =  factor(mydat.P$ld50_method, 
                       levels = c("sc","iv","ip","im")), 
dim = mydat.P$dimensions, 
egg.bin = mydat.P$egg.bin, 
env = factor(mydat.P$env, 
             levels = c("terrestrial","arboreal","aquatic")),
con = mydat.P$con,
fam = mydat.P$fam,
mod_class_p = mydat.P$mod_class,
prey_mean = prey_mean)

```

##Prey model
Run the main model with prey size included 

```{r prey model run}
#if you want to include species and ld50 method as a fixed term.
prior<-list(R = list(V = diag(2), nu=0.002), 
            G = list(G1=list(V = diag(2)/2,n = 2, 
                             alpha.mu=rep(0,2), 
                             alpha.V=diag(2)*10^3),
                     G2=list(V = diag(2)/2,n = 2, 
                             alpha.mu=rep(0,2), 
                             alpha.V=diag(2)*10^3)))


nitt_p <- 120000
thin_p <- 50
burnin_p <- 20000


######with prey included
prey_mod_long <- MCMCglmm(cbind(vol.v,ld50) ~ 
                            trait:mass.aa + 
                            trait:ld50_method + 
                            trait:prey_mean + 
                            trait:dim + 
                            trait:egg.bin + 
                            trait:weigted.dist.v, 
                          random = ~us(trait):animal + 
                                    us(trait):species,
                          rcov=~us(trait):units, 
                          data = mydat.P, 
                          pedigree = prey.tree, 
                          prior = prior,
                          family=c("gaussian","gaussian"), 
                          nitt = nitt_p, thin = thin_p, burnin = burnin_p,
                          verbose = FALSE)



ld_50_prey_phlyo <- prey_mod_long$VCV[,"traitld50:traitld50.animal"]/
                    (prey_mod_long$VCV[,"traitld50:traitld50.animal"] +
                     prey_mod_long$VCV[,"traitld50:traitld50.species"] +
                     prey_mod_long$VCV[,"traitld50:traitld50.units"])

ld_50_prey_species <- prey_mod_long$VCV[,"traitld50:traitld50.species"]/
                    (prey_mod_long$VCV[,"traitld50:traitld50.animal"] +
                     prey_mod_long$VCV[,"traitld50:traitld50.species"] +
                     prey_mod_long$VCV[,"traitld50:traitld50.units"])

ld_50_prey_units <- prey_mod_long$VCV[,"traitld50:traitld50.units"]/
                    (prey_mod_long$VCV[,"traitld50:traitld50.animal"] +
                     prey_mod_long$VCV[,"traitld50:traitld50.species"] +
                     prey_mod_long$VCV[,"traitld50:traitld50.units"])

hdr(ld_50_prey_phlyo)
hdr(ld_50_prey_species)
hdr(ld_50_prey_units)



vol_prey_phlyo <-  prey_mod_long$VCV[,"traitvol.v:traitvol.v.animal"]/
                    (prey_mod_long$VCV[,"traitvol.v:traitvol.v.animal"] +
                     prey_mod_long$VCV[,"traitvol.v:traitvol.v.species"] +
                     prey_mod_long$VCV[,"traitvol.v:traitvol.v.units"])

vol_prey_spec <-  prey_mod_long$VCV[,"traitld50:traitvol.v.species"]/
                    (prey_mod_long$VCV[,"traitvol.v:traitvol.v.animal"] +
                     prey_mod_long$VCV[,"traitvol.v:traitvol.v.species"] +
                     prey_mod_long$VCV[,"traitvol.v:traitvol.v.units"])

vol_prey_units <-  prey_mod_long$VCV[,"traitvol.v:traitvol.v.units"]/
                    (prey_mod_long$VCV[,"traitvol.v:traitvol.v.animal"] +
                     prey_mod_long$VCV[,"traitvol.v:traitvol.v.species"] +
                     prey_mod_long$VCV[,"traitvol.v:traitvol.v.units"])


hdr(vol_prey_phlyo)
hdr(vol_prey_spec)
hdr(vol_prey_units)



summary(prey_mod_long)
```


##Predator-prey size allometry
Model of predator prey allometry.

```{r pred-prey model}
#if you want to include species and ld50 method as a fixed term.
prior2<-list(R = list(V = 1, nu=0.002), 
             G = list(G1=list(V = 1,n = 2, 
                              alpha.mu= 0, 
                              alpha.V = 10^3)))


######with prey included
prey_mass_mod <- MCMCglmm(prey_mean ~ 
                            mass.aa, 
                          random = ~ animal,
                          rcov=~units, 
                          data = mydat.P, 
                          pedigree = prey.tree, 
                          prior = prior2, 
                          family=c("gaussian"), 
                          nitt = 600000, thin =250, burnin = 100000,
                          verbose = FALSE)

summary(prey_mass_mod)

```
